---
title: "Project 2: a study on the ToothGrowth database"
output: pdf_document
---

# 0) Introduction

In this short note, we will do a simple statistical analysis of the ToothGrowth
database contained in the datasets library. Note for the reviewers: since it
is quite useless to go back and forth between the code put in appendix and 
the core of the project, I merged the two.

# I) Exploratory analysis

We start by loading the datasets library

```{r,results="hide"}
library(datasets)
 dim(ToothGrowth)
```

we see that ToothGrowth is a $60\times 3$ dataframe. The first column
(using `?ToothGrowth`)
corresponds to the length of teeth in each of 10 guinea pigs, the second column
to two different delivery methods (orange juice or ascorbic acid). The thirs
column corresponds to three different dose levels of Vitamin 
C (0.5, 1, and 2 mg). Since the Vitamine dose and the delivery method are qualitative variables, we 
will convert them into factors

```{r}
ToothGrowth$supp<-as.factor(ToothGrowth$supp)
 ToothGrowth$dose<-as.factor(ToothGrowth$dose)
```

Let us consider following the exploratory graph


```{r,fig.height=3.7,fig.width=6}
 library(lattice)
histogram(~len|dose*supp,data=ToothGrowth, xlab="teeth length",
    main=list(label="teeth length as a function of Vitamine dose and delivery method"
              ,cex=0.7),col="cornflowerblue")
```

and  a summary of the data frame gives

```{r}
summary(ToothGrowth)
```

The histograms represent the length of the teeth as a function of both the
delivery method and the vitamine dose. As one can see, as the Vitamine dose 
increases, the teeth length seems to increas as well. It also seems that the 
Orange Juice delivery method leads to slightly longer teeth than with 
ascorbic acid. This will be our working hypotheses for the next section.

# I) Does the teeth length depends on the delivery method?


Firstly, does the teeth length depends on the delivery method? We can compute 
the mean and the standard deviation of both classes with
```{r}
meanOJ<-mean(ToothGrowth$len[ToothGrowth$supp=="OJ"])
meanVC<-mean(ToothGrowth$len[ToothGrowth$supp=="VC"])
sdOJ<-sd(ToothGrowth$len[ToothGrowth$supp=="OJ"])
sdVC<-sd(ToothGrowth$len[ToothGrowth$supp=="VC"])
```

We will assume equal variance. Our hypothesis $H_0$ is : the teeth length is the
same whatever the delivery method, while our $H_a$ hypothesis is that
the OJ delivery method gives different teeth length than the VC one. Given the small size 
of our sample, we will do a confidence interval analysis with the quantile of the
student-t distribution. We will assume equal variance. The relevant formulas are
\begin{align}
[\text{conf int}]=\null& \mu_{_{\rm OJ}}-\mu_{_{\rm VC}}\pm
qt(0.975,N_{_{\rm OJ}}+N_{_{\rm VC}}-2)\times \Sigma\times
\sqrt{\frac{1}{N_{_{\rm OJ}}}+\frac{1}{N_{_{\rm VC}} }}
\end{align}
where the pooled variance is
\begin{align}
\Sigma=\sqrt{\frac{(N_{_{\rm OJ}}-1)\sigma^2_{_{\rm OJ}}
+(N_{_{\rm VC}}-1)\sigma^2_{VC}}{N_{_{\rm OJ}}+N_{_{\rm VC}}-2}}
\end{align}

the R calculation gives 
```{r}
NOJ<-sum(ToothGrowth$supp=="OJ")
NVC<-sum(ToothGrowth$supp=="VC")
Sigma<-sqrt(((NOJ-1)*sdOJ^2+(NVC-1)*sdVC^2)/(NOJ+NVC-2))
meanOJ-meanVC+c(-1,1)*qt(0.975,NOJ+NVC-2)*Sigma*sqrt(1/NOJ+1/NVC)
```

the same result can be obtained with
```{r}
t.test(ToothGrowth$len[ToothGrowth$supp=="OJ"],
       ToothGrowth$len[ToothGrowth$supp=="VC"],paired=FALSE,var.equal=TRUE)$conf
```

Since $0$ is in the $95\%$ confidence interval, we \textbf{fail to reject the 
null Hypothesis $H_0$}. 

# I) Does the teeth length depends on the vitamine dose ?


Secondly, does the teeth length depends on the vitamine dose? This time we will
just use the built-in R function. We will assume equal variance. 
Our hypothesis $H_0$ is : the mean of the length
teeth does not depend on the vitamine dose, while our $H_a$ hypothesis is that
the vitamine dose affects the teeth length. Since the t-test is 
designed to compare between two groups, we will do three t-test: between the 
$0.5$ and the $1$mg doses, the $0.5$ and the $2$mg doses and finally between the
$1$ and the $2$mg doses.

1. $0.5$ vs $1$mg dose:
```{r}
t.test(ToothGrowth$len[ToothGrowth$dose=="1"],
       ToothGrowth$len[ToothGrowth$dose=="0.5"],paired=FALSE,var.equal=TRUE)$conf
```

2. $0.5$ vs $2$mg dose:
```{r}
t.test(ToothGrowth$len[ToothGrowth$dose=="2"],
       ToothGrowth$len[ToothGrowth$dose=="0.5"],paired=FALSE,var.equal=TRUE)$conf
```

3. $1$ vs $2$mg dose:
```{r}
t.test(ToothGrowth$len[ToothGrowth$dose=="2"],
       ToothGrowth$len[ToothGrowth$dose=="1"],paired=FALSE,var.equal=TRUE)$conf
```
In all cases, $0$ is in not in the $95\%$ confidence interval and we \textbf{reject the 
null Hypotheses $H_0$}. 